Automated topic naming to support analysis of software maintenance activities

### 2010-08-20 icse/icse2011.tex
Updated intro with reference to NFRs and ISO

% why are topics useful
Often developers interact with abstractions of the artifacts from which topics are derived, rather than the artifact itself. 
Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
Topics, too, are a valuable and automatically extractable abstraction of software artifacts.
Topics abstracts these artifacts by their subject matter, rather than their concrete representation. 
Much like Aspect-Oriented Programming, topic labeling can identify cross-cutting concerns in a software project~\cite{Baldi2008}.
The topic of development is relevant because it identifies the distinct issues facing development at a certain time.

The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
% get neil's data mining point in there
To date, the results of topic analysis and concept location are project specific, and not generalizable.
% get the interpretation aspect
Much of the current research stops short of interpreting the purpose of the topic. 
That is, extracted topics are usually manually named and labelled by those who read the topic words.
We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.

% topics are numerous?
Deriving such topics is difficult. 
Topic analysis produces raw topics that are represented as unlabelled context-free word-lists or word distributions.
Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
This manual approach is impractical when one has to handle hundreds of different topics. To scale to meaningfully sized projects, we need automatic assistance to determine the topic label.

@@

% harp on data mining more
In general the data mining of software artifacts tends to be very project specific, yet NFRs are not. 
NFRs are prevalent in almost all software systems.
% In this paper we will show that NFR related topics occur quite frequently.
This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. In particular, we use software quality models to generalize topics across projects. 
This allows us to data mine software systems with the intention of comparing their NFR-related development topics.
% in order to achieve generalizable, cross project data mining results.

@@

%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
\begin{figure*}
  \centering
 \includegraphics[width=\textwidth]{figures/unsupervised-bar}
 \caption{ROC values for each NFR, per project and wordlist.}
%\todo[inline]{this is too tiny in print}
  \label{fig:maxdb-unsup-results}
\end{figure*}

>>>>

% why are topics useful
Developers often interact with abstractions of the artifacts from which topics are derived, rather than the artifact itself. 
Such abstractions include project dashboards~\cite{kersten2005mylar}, tags~\cite{treude2010}, and complexity metrics \cite{mccabe1976complexity}. 
Topics, too, are a valuable and automatically extractable abstraction of software artifacts.
Topics abstracts these artifacts by their subject matter, rather than their concrete representation. 
Much like aspect-oriented programming, topic labeling can identify cross-cutting concerns in a software project (as in~\cite{Baldi2008}).
The topic of development is relevant because it identifies the distinct issues facing development at a particular time.

The goal of concept location and topic analysis research is to recover these topics and the related software artifacts.
% get neil's data mining point in there
To date, the results of topic analysis and concept location are project specific, and not generalizable.
% get the interpretation aspect
Much of the current research stops short of interpreting the purpose of the topic. 
That is, extracted topics are usually manually named and labelled by those who read the topic words.

% topics are numerous?
Deriving such topics is difficult. 
Topic analysis produces% raw topics that are represented as
 unlabelled context-free word-lists or word distributions.
Topics have to be identified by interpreting the prevalent words in a word distribution and by inspecting related documents. 
This manual approach is impractical when one has to handle hundreds of different topics. To scale to meaningfully sized projects, we need automatic assistance to determine the topic label.
We propose a method of interpreting and labelling topics, and their related artifacts, that is generalizable and relevant to multiple projects.

@@

% harp on data mining more
In general, the mining of software artifacts tends to be very project specific, yet NFRs are not. There is a series of standards on NFRs (such as \cite{iso9126}) specifically intended to apply to projects of varying types (indicating this is possible, if perhaps not universally accepted).
% NFRs are prevalent in almost all software systems.
% In this paper we will show that NFR related topics occur quite frequently.
This informs our method of generalizing the data mining of these software artifacts to leverage the global domain knowledge of software engineering, specifically NFRs. In particular, we use software quality models to generalize topics across projects. 
This allows us to mine software systems with the intention of comparing their NFR-related development topics.
% in order to achieve generalizable, cross project data mining results.

@@

%\todo[inline]{since ROC is new-sounding, I wonder if a more lengthy explanation would be merited. An obvious question is why not precision/recall? -- ROC is hot right now and it combines both}
% Rebuttal: ROC is not new and it is what the machine learning community, search based software engineering and now the MSR community are using +  it has natural meaningful quality mappings.
\begin{figure*}
  \centering
 \includegraphics[width=\textwidth]{figures/unsupervised-bar}
 \caption{Performance, ROC values, of Unsupervised topic labelling for
   each NFR and per word-list. This graph shows how well the
   unsupervised topic labelling matched our manual annotations.}

%\todo[inline]{this is too tiny in print}
  \label{fig:maxdb-unsup-results}
\end{figure*}
